---
title: "RA Luis Zapata - PHD Luca - Anderson School of Economics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Explanation of Code

I will be utilizing matched employer-employee data from Germany. In this [link](https://fdz.iab.de/en/int_bd_pd/the-sample-of-integrated-employer-employee-data-sieed-sieed-7518-version-1/), you can find a description of the dataset. Specifically, there is a file at the link containing a detailed description of the data ([this](https://doku.iab.de/fdz/reporte/2020/DR_14-20_EN.pdf)) and the test data, which is a fake data. Unfortunately, access to the real data is restricted due to confidentiality constraints associated with the dataset.

This [paper](https://davidcard.berkeley.edu/papers/QJE-2013.pdf) is an excellent piece that utilizes the same dataset, implementing certain methodologies and approaches that I aim to replicate or use as a foundation for my analysis. Additionally, the paper discusses some issues with the dataset, including the fact that wages are capped. When wages exceed a specific threshold, they are recorded as the threshold value. The authors employ a methodology to estimate the actual wage beyond the threshold, and I plan to adopt a similar approach. The initial task will involve replicating their code, translating it into either R. 

## Preparing the Data

We beggin by calling the libraries as follow:

```{r, message = FALSE,warning = FALSE}
library(data.table) #Table data handler
library(censReg)  # For Tobit regression
library(dplyr) #Also data handler
library(tidyverse,quietly = TRUE) #For data managment
library(lubridate)  # For easy date handling
library(haven) # For reading Stata files
```

## Call the data

We call the Stata data (.dta) from our folder and store it in a list. 
```{r}
dta_files = list.files("SIEED_7518_v1_test", pattern = ".dta")
dta_list = lapply(dta_files, function(x) read_dta(paste0("SIEED_7518_v1_test/", x)))
print(class(dta_list))
```

I now print the first row of all the dataframes in the list to get a sense of the data. 
```{r}
for (i in 1:length(dta_list)){
  print(dta_files[i])
  print(head(dta_list[[i]],1))
}
```

