---
title: "RA Luis Zapata - PHD Luca - Anderson School of Economics"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Explanation of Code

I will be utilizing matched employer-employee data from Germany. In this [link](https://fdz.iab.de/en/int_bd_pd/the-sample-of-integrated-employer-employee-data-sieed-sieed-7518-version-1/), you can find a description of the dataset. Specifically, there is a file at the link containing a detailed description of the data ([this](https://doku.iab.de/fdz/reporte/2020/DR_14-20_EN.pdf)) and the test data, which is a fake data. Unfortunately, access to the real data is restricted due to confidentiality constraints associated with the dataset.

This [paper](https://davidcard.berkeley.edu/papers/QJE-2013.pdf) is an excellent piece that utilizes the same dataset, implementing certain methodologies and approaches that I aim to replicate or use as a foundation for my analysis. Additionally, the paper discusses some issues with the dataset, including the fact that wages are capped. When wages exceed a specific threshold, they are recorded as the threshold value. The authors employ a methodology to estimate the actual wage beyond the threshold, and I plan to adopt a similar approach. The initial task will involve replicating their code, translating it into either R. 

## Data Sets managment
The documentation containts information about the datasets that contain yearly data and those that involve worker flows and establishment dynamics such as entries, exits, inflows, and outflows. Here is a breakdown:

1. **Datasets with Years in Their Name**: These datasets are annual files that contain variable blocks with data from each year from 1975 to 2018. They are named using the format `SIEED_7518_v1_bhp_v1_jjjj.dta`, where `jjjj` represents the specific year (e.g., `1975`, `2016`, etc.). Each file corresponds to a particular year and includes data up to the reference date of June 30th for that year. These files provide data specific to each year and can include both individual and firm-level information, focusing on the characteristics and dynamics relevant to that particular year.

  - **Basis Establishment File (`SIEED_7518_v1_bhp_basis_v1.dta`)**: Contains aggregated information of the estabilishment place of work and economic activities. Contains the establishment number, the year, and aggregated variables about the place of work and economic activities as of the reference date, 30 June. It is meant to be linked with the Individual File for comprehensive analysis.
   
   - **Individual-level File (`SIEED_7518_v1_bhp_basis_v1.dta`)**: Contain comprehensive individual-level data, including information about employment spells, wages, demographics, and possibly links to establishments where the individuals worked. Researchers can use SIEED_7518_v1.dta as a starting point for their analysis and then merge it with other specific datasets like the SIEED_7518_v1_bhp_basis_v1.dta or year-specific files to add layers of context or to focus on particular aspects of the labor market.

2. **Datasets with Entry, Exit, Inflow, and Outflow in Their Names**: These datasets specifically contain information about the dynamics of the workforce within establishments:

   - **Entries (`SIEED_7518_v1_bhp_entry_v1.dta`)**: This file contains data about the entries into establishments, which could be indicative of the creation of new positions or filling of existing ones.

   - **Exits (`SIEED_7518_v1_bhp_exit_v1.dta`)**: This file includes data on exits from establishments, which might reflect layoffs, terminations, or retirements.
   
   - **Inflows (`SIEED_7518_v1_bhp_inflow_v1.dta`)**: This dataset captures the inflow of employees to establishments, providing insights into hiring trends.
   
   - **Outflows (`SIEED_7518_v1_bhp_outflow_v1.dta`)**: Conversely, this dataset records the outflow of employees from establishments, which could inform on turnover rates and employment stability.

These datasets can be accessed and used following an approved application for research purposes and are essential for analyzing employment trends and patterns within establishments over time. The merging of individual-level and establishment-level data within statistical software such as Stata is necessary to perform comprehensive analyses. The datasets are meant to be linked via identifiers like the establishment number and the reference year.

## Structure of the Data

The documentation describe the dataset subsets and how to use them, particularly in the context of the Sample of Integrated Employer-Employee Data (SIEED). These subsets are designed to be used in conjunction with one another and are linked via specific identifiers such as establishment numbers and individual IDs.

Here's a brief overview of the structure and use of the dataset subsets as explained in the PDF:

1. **Basis Establishment File**: Contains the establishment number, the year, and aggregated variables about the place of work and economic activities as of the reference date, 30 June. It is meant to be linked with the Individual File for comprehensive analysis.

2. **Individual File**: Includes personal identifiers, personal variables, information on employment, and place of residence, among other details.

3. **Episode Splitting**: The process of dealing with overlapping observations within an account by creating artificial observations with new dates, leading to non-overlapping periods.

4. **Sampling Procedure**: Outlines the steps taken to select a sample of establishments and individuals for inclusion in the SIEED.

5. **Data Use**: Describes the conditions under which the SIEED data may be accessed and analyzed, including research visits and remote data access following application approval by the FDZ.

6. **File Architecture**: Discusses how data are stored in separate files for individual and establishment-related information, and that additional files with more detailed establishment information can be provided upon justified application.

To use the data, researchers are advised to **merge the individual-level and establishment-level** data within their statistical software. 

It is suggested to merge the datasets based on establishment numbers and the year, which should correspond between the individual and establishment files.

Each subset of the SIEED data has a specific role and, when combined, can provide a comprehensive view of employment biographies and establishment characteristics. 

For more inquiries about the data, go to the documentation link provided above.

## Notations about the data.

### Dates
The data in this file are not longitudinal (spell data) but rather give information that is only valid precisely on June 30th of each year. 
In the context of the SIEED data, there's a distinction between the original start/end dates and the episode start/end dates:

- **Original Start/End Dates (`begorig` and `endorig`)**: These dates correspond to the original start and end dates of the notification for an employment spell as recorded in the administrative data. They represent the formal notification dates of employment relationships as they were initially reported. **An "employment spell" refers to a continuous period during which an individual is employed at a particular job. It starts on the day the individual begins that job and ends on the day the job terminates.** This period could encompass various employment statuses, but it's typically defined by uninterrupted service with a single employer.

- **Episode Start/End Dates (`begepi` and `endepi`)**: These dates are generated after the process of episode splitting. Episode splitting is used to deal with overlapping observations within an account by creating non-overlapping periods. **Overlapping spells occur when the dataset shows that an individual has two or more jobs at the same time, which may be due to multiple job holdings or errors in data recording.** Episode splitting resolves this by dividing the time into discrete, non-overlapping episodes. Each episode cor    responds to a unique job, so that no two episodes overlap in time. Thus, the episode dates reflect the start and end of these non-overlapping periods after splitting.

Here’s how the splitting is done:

- **Parallel Periods:** Overlapping periods are identified, and parallel periods within the same time frame are created.
- **Splitting Into Non-overlapping Intervals:** These parallel periods are then adjusted to create intervals that do not overlap. Each period is assigned a unique start and end date that fits within the original period of employment without intersecting with another job spell.
- **Generated Variables:** New variables, such as the episode start (begepi) and episode end (endepi) dates, are generated to mark the beginning and end of these non-overlapping periods.


The episode dates will always be within the range of the original dates, with the episode start date (`begepi`) being equal to or later than the original start date (`begorig`), and the episode end date (`endepi`) being equal to or earlier than the original end date (`endorig`). This adjustment is crucial for analyses that require the use of spell data without any overlaps.

### Wages
The dataset reports wages in euros, reflecting gross wages subject to social security contributions, not adjusted for inflation. Here's a concise overview:

- **Wage Reporting**: Until 1998, only wages within specific thresholds were recorded. Post-1998, reporting expanded to include lower wages.
- **Gross Daily Wage**: Calculated by dividing fixed-period wages by the number of days in the period, presenting wages on a daily basis.
- **Components of Gross Wage**:
  1. **Basic Salary**: Core earnings from salary agreements or hourly wages.
  2. **Overtime Pay**: Compensation for hours beyond the regular work schedule.
  3. **Bonuses and Commissions**: Earnings from performance or company success.
  4. **Allowances**: Payments for travel, meals, housing, etc.
  5. **Other Earnings**: Additional income like stock options or profit sharing.

Gross wage is the total income before any deductions such as taxes or social security contributions.

The marginal part-time income threshold in the dataset refers to the minimum earnings level below which employment is considered marginal or "mini-job," and thus may not be fully subject to social security contributions. The upper earnings limit, or contribution assessment ceiling, is the maximum income level up to which social security contributions are assessed. These thresholds can vary by year and between eastern and western Germany.

- **Marginal Part-Time Income Threshold**: Below this level, employment is marginally part-time, often exempt from full social security contributions. Since April 1999, earnings below this threshold have been recorded in the dataset.
  
- **Contribution Assessment Ceiling**: This is the maximum limit of earnings that are subject to social security contributions. Earnings above this ceiling are not subject to additional contributions, and the dataset includes wages up to this limit.

## Preparing the Data

We beggin by calling the libraries as follow:

```{r, message = FALSE,warning = FALSE}
library(data.table) #Table data handler
library(censReg)  # For Tobit regression
library(dplyr) #Also data handler
library(tidyverse,quietly = TRUE) #For data managment
library(lubridate)  # For easy date handling
library(haven) # For reading Stata files
```

## Call the data

We call the Stata data (.dta) from our folder and store it in a list. 
```{r}
dta_files = list.files("SIEED_7518_v1_test", pattern = ".dta")
dta_list = lapply(dta_files, function(x) read_dta(paste0("SIEED_7518_v1_test/", x)))
print(class(dta_list))
```

I now print the last 10 rows of the list to see what kind of data we are dealing with. 
```{r}
print(tail(dta_files, 10))
```

We can see that the data have one kind related with years, while the others have the names: basis, entry, exit, infow and outflow associated with them.
 

```{r}
#On dta_files there are data with year on the name, and other without it.
#Lets filter the ones with year on the name

dta_files_2 <- dta_files[str_detect(dta_files, "^(SIEED_7518_v1_bhp_20|SIEED_7518_v1_bhp_19)")]

dta_files_3 <- dta_files[!dta_files %in% dta_files_2]

print(dta_files_3)
```
We know merge all the data that are related with the yeards 2019 and 2020. 

```{r}
dta_year = tibble("Address" = dta_files_2) %>%
  mutate("Year" = str_sub(Address, start = 19, end = 22))
#We extract the year, wich is the second four digits of the file name
#Lets read all the dta in the Address column and store them in a list. Lets name the list with the year of the data
dta_list_2 = lapply(dta_files_2, function(x) read_dta(paste0("SIEED_7518_v1_test/", x)))
names(dta_list_2) = dta_year$Year
#Lets merge all the data in dta_list_2, in order to create one big tibble. We also create one column with name "Year" to store the year of the data
dta_list_2 = dta_list_2 %>%
  bind_rows(.id = "Year") %>%
  select(-Year)
print(paste("The size of the data is: rows:", nrow(dta_list_2), " and columns:", ncol(dta_list_2)))
print(head(dta_list_2,3))
```


```{r}
print(glimpse(dta_list_2))
```

We still have to understand better the way the data is gathered. We have to read the documentation in order to do it.

```{r}
print(dta_files_3)
```